---
title: "L06 Annotation & Positioning"
author: "Anna Wagman"
subtitle: Data Visualization (STAT 302)
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
    highlight: tango
---

```{r global_options}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, comment = FALSE, dpi = 300)
```

## Overview

The goal of this lab is to explore methods for annotating and positioning with `ggplot2` plots. This lab also utilizes `scale_*` to a greater degree which is part of our next reading. In fact, students may find going through/reading chapter [11 Colour scales and legends](https://ggplot2-book.org/scale-colour.html) useful.

## Datasets

We'll be using the `blue_jays.rda`, `titanic.rda`, `Aus_athletes.rda`, and `tech_stocks.rda` datasets.

<br>

```{r load-packages-data}
# Load package(s)
library(tidyverse)
library(skimr)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(lubridate)
library(patchwork)
library(scales)

# Load datasets
load(file = "data/blue_jays.rda")
load(file = "data/titanic.rda")
load(file = "data/Aus_athletes.rda")
load(file = "data/tech_stocks.rda")
load(file = "data/corruption.rda")



```

<br>

## Exercises

Complete the following exercises.

<br>

### Exercise 1

Using `blue_jays.rda` dataset, recreate the following graphic as precisely as possible.

*Hints:*

- Transparency is 0.8
- Point size 2
- Create a `label_info` dataset that is a subset of original data, just with the 2 birds to be labeled
- Shift label text horizontally by 0.5
- See 8.3 building custom annotations 
- Annotation size is 4
- Classic theme


```{r}
label_info_ex1 <- blue_jays %>% #subset of original data
  select(Mass, Head, KnownSex) %>%
  arrange(desc(Head)) %>% #arrange by head size
  slice(n = c(1, 28))

caption <- 'Head length versus body mass for 123 blue jays'

xrange <- range(blue_jays$Mass)
yrange <- range(blue_jays$Head)

##plot label_info

ggplot(data = blue_jays,
       aes(x = Mass, y = Head, color = KnownSex)) +
  
  geom_point(size = 2, alpha = 0.8) + #given point size = 2 and transparency = .8
  labs(
    x = 'Body mass (g)',
    y = 'Head length (mm)') +
  
  theme_classic() + #theme is classic
  geom_text(data = label_info_ex1, label = c('M', 'F'), 
            nudge_x = 0.5) + #shift horiz by .5
  annotate(geom = "text", x = xrange[1], y = yrange[2],
           label = caption, hjust = 0, vjust = 1, size = 4) +
  theme(legend.position = 'none')
  
  

```


<br>

### Exercise 2

Using `tech_stocks` dataset, recreate the following graphics as precisely as possible.

<br>

#### Plot 1

*Hints:*

- Create a `label_info` dataset that is a subset of original data, just containing the last day's information for each of the 4 stocks
- serif font
- Annotation size is 4


PLOT 1:

```{r}
label_info_ex2 <- tech_stocks %>%
  group_by(ticker) %>%
  filter(date == max(date))

caption_ex2 <- 'Stock price over time for four major tech companies'

#x and y range
xrange2 <- range(tech_stocks$date) 
yrange2 <- range(tech_stocks$price_indexed)

tech_stocks %>%
  
#plot subset of data  
  ggplot(aes(x = date, y = price_indexed)) +
  geom_line(aes(color = company)) +
  labs(x = NULL, y = 'Stock price, indexed') +
  geom_text(data = label_info_ex2, label = label_info_ex2$company) +
  annotate(geom = 'text', x = ymd('2006-05-01'), y = yrange2[2],
           label = caption_ex2, hjust = 0, vjust = 1, size = 4, family = 'serif') +
  
  theme_minimal() +
  theme(legend.position = 'none')
  
  
```



<br>

#### Plot 2

*Hints:*

- Package `ggrepel`
- Annotation size is 4
- `box.padding` is 0.6
- Minimum segment length is 0 
- Horizontal justification is to the right
- seed of 9876

```{r}
tech_stocks %>%
  
  ggplot(aes(x = date, y = price_indexed)) +
  geom_line(aes(color = company)) +
  labs(x = NULL, y = 'Stock price, indexed') +
  theme_minimal() +
  theme(legend.position = 'none') +
  
  annotate(geom = 'text', x = ymd('2006-05-01'), y = yrange2[2],
           label = caption_ex2, hjust = 0, vjust = 1, size = 4, family = 'serif') +
  
  ##USE GGREPEL PACKAGE
  
  geom_text_repel(data = label_info_ex2,
                  aes(label = company),
                  hjust = 'right',
                  box.padding = 0.6,
                  min.segment.length = 0,
                  seed = 9876)
```



<br>

### Exercise 3

Using the `titanic.rda` dataset, recreate the following graphic as precisely as possible.



*Hints:*

- Create new variable for that uses `died` and `survived` as levels/categories
- Hex colors: `#D55E00D0`, `#0072B2D0` (no alpha is being used)


```{r}
ggplot(titanic, aes(sex, fill = sex)) +
  geom_bar() +
  facet_grid(
    factor(survived, labels = c("died", "survived"))
    ~
    class
  ) +
  scale_fill_manual(values = c("#D55E00D0", "#0072B2D0")) +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}

```


<br>

### Exercise 4

Use the `athletes_dat` dataset --- extracted from `Aus_althetes.rda` --- to recreate the following graphic as precisely as possible. Create the graphic twice: once using `patchwork` and once using `cowplot`.


```{r}
# Get list of sports played by BOTH sexes
both_sports <- Aus_athletes %>%
  distinct(sex, sport) %>%
  count(sport) %>%
  filter(n == 2) %>%
  pull(sport)

# Process data
athletes_dat <- Aus_athletes %>%
  filter(sport %in% both_sports) %>%
  mutate(sport = case_when(
    sport == "track (400m)" ~ "track",
    sport == "track (sprint)" ~ "track",
    TRUE ~ sport
  ))
```

*Hints:*

- Build each plot separately
- Hex values for shading: `#D55E0040` and `#0072B240` (bottom plot),
`#D55E00D0` & `#0072B2D0` (for top two plots) --- no `alpha`
- Hex values for outline of boxplots: `#D55E00` and `#0072B2`
- Boxplots should be made narrower; 0.5
- Legend is in top-right corner of bottom plot
- Legend shading matches hex values for top two plots
- Bar plot lower limit 0, upper limit 95; shift bar labels by 5 units and top justify; label size is 5
- `rcc`: red blood cell count; `wcc`: white blood cell count
- Size 3 will be useful


#### Using `patchwork`

```{r}
##number on bar plot
bar_numb <- athletes_dat %>%
  count(sex)

plotA <- ggplot(athletes_dat, aes(sex, fill = sex)) + 
  geom_bar(show.legend = FALSE)+
  scale_fill_manual(values = c("#D55E00D0", "#0072B2D0")) +
  scale_x_discrete(
      name = NULL
    , labels = c("female", "male")
  ) +
  scale_y_continuous(
      name = "number"
    , breaks = seq(0, 100, 25)
    , limits = c(0, 95)
  ) + 
  theme_minimal() +
  geom_text(data = bar_numb, aes(y = n, label = n), size = 5, nudge_y = -5, vjust = 'top')
```

```{r}
plotB <- ggplot(athletes_dat, aes(rcc, wcc)) +
  geom_point(
      aes(fill = sex)
    , shape = 21
    , color = "white"
    , size = 3
    , show.legend = FALSE
    ) + 
  scale_fill_manual(values = c("#D55E00D0", "#0072B2D0")) + 
  scale_x_continuous(name = "RBC Count") + 
  scale_y_continuous(name = "WBC count") + 
  theme_minimal()
```

```{r}
sex_labs <- c("female", "male")

plotC <- ggplot(athletes_dat, aes(sport, pcBfat)) +
  geom_boxplot(
      aes(color = sex, fill = sex)
    , width = 0.5
    ) +
  scale_fill_manual(
      name = NULL
    , labels = sex_labs 
    , values = c("#D55E0040", "#0072B240")) + 
  scale_color_manual(
      name = NULL
    , labels = sex_labs
    , values = c("#D55E00D0", "#0072B2D0")) +
  guides(fill = guide_legend(
      ncol = 2
    , override.aes = list(
        fill = c("#D55E00D0", "#0072B2D0")
      , color = "transparent"
      )
    )
  ) +
  xlab(NULL) + 
  ylab("% body fat") + 
  theme_minimal() + 
  theme(
      legend.position = c(1, 1)
    , legend.justification = c(1, 1)
    #set top legend to zero
    , legend.margin = margin(t = 0)
  )
```

##USE PATCHWORK:
```{r}
(plotA + plotB) / plotC + plot_layout()
```


<br>



<br>



<br>

#### Using `cowplot`

Use [`cowplot::plot_grid()`](https://wilkelab.org/cowplot/articles/plot_grid.html) to combine them.

<br>

### Exercise 5

Create the following graphic using `patchwork`.

*Hints:*

- Use plots created in exercise 4
- Useful values: 0, 0.45, 0.75, 1
- inset theme is classic


```{r}
plotB +
  inset_element(plotA + theme_classic(),
                left = 0.75, bottom = 0, right = 1, top = 0.45) +
  
  plot_annotation(tag_levels = c('A', 'B'))
```



<br>


