---
title: "L05 Maps"
author: "Anna Wagman"
subtitle: Data Visualization (STAT 302)
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
    highlight: tango
---

## Overview

The goal of this lab is to explore various was of building maps with `ggplot2`.

Challenges are not mandatory for students to complete. We highly recommend students attempt them though. *We would expect graduate students to attempt the challenges.*

## Datasets

We'll be using the `US_income.rda` dataset which is already in the `/data` subdirectory in our **data_vis_labs** project. You'll also be downloading your own data to build maps.

```{r load-packages, warning=FALSE, message=FALSE}
# Load package(s)
library(tidyverse)
library(skimr)
library(maps)

# Load dataset(s)
load(file = "data/US_income.rda")

```

## Exercises

### Exercise 1

#### Plot 1

Make a county map of a US state using `geom_polygon()`. Maybe use your home state (mine is below) or a favorite state. Please do not use the state in example provided in the book. 

Optional: Consider adding major cities (or your home town).

*Hints:*

-   See section [6.1](https://ggplot2-book.org/maps.html#polygonmaps) in our book.
-   Void theme

```{r}
#select my home state 
ny <- map_data("county", "new york") %>% 
  select(long, lat, group, id = subregion)

#plot state, divide by counties
ggplot(ny, aes(x = long, y = lat)) +
  #title
  ggtitle("New York") +
  #geompoly method
  geom_polygon(aes(group = group), fill = "white" , color = "grey35") +
  #size
  coord_quickmap() +
  #theme
  theme_void()
```




<br>

#### Plot 2

Now use `geom_sf()` instead. You'll need to download data for this. You can use either the `tigris` ([github page](https://github.com/walkerke/tigris)) or `raster` packages. Either `tigris`' `counties()` with `cb = TRUE` or `raster`'s `getData()` could be useful.

```{r}
library(raster)
library(tigris)
library(sf)
library(ggplot2)
library(dplyr)

##create same plot using geom_sf

##use tigris (cb = TRUE) and class = sf
#isolate homestate new york
nyny <- counties(state = "NY", cb = TRUE, class = "sf")

##plot new york using geom_sf
ggplot(data = nyny, aes(geometry = geometry)) + 
  geom_sf(fill = "white") + 
  coord_sf() + 
  theme_void() +
  ggtitle("New York")





```



<br>

### Exercise 2

Using `US_income` dataset, recreate the following graphics as precisely as possible.

<br>

```{r}
# Setting income levels
US_income <- mutate(
  US_income,
  income_bins = cut(
    ifelse(is.na(median_income), 25000, median_income),
    breaks = c(0, 40000, 50000, 60000, 70000, 80000),
    labels = c("< $40k", "$40k to $50k", 
               "$50k to $60k", "$60k to $70k", "> $70k"),
    right = FALSE
  )
)
```

<br>

##### Plot 1

*Hints:*

-   `geom_sf()` --- boundary color is `"grey80"` and size is 0.2
-   `viridis` package (`discrete = TRUE` in `scale_*` function)
-   Void theme


```{r}
library(viridis)


ggplot(data = US_income) +
  #use geom_sf
  geom_sf(aes(geometry = geometry, fill = income_bins)
    #set boundary color = grey80
    , color = "grey80"
    #set size = 0.2
    , size = 0.2) +
  
  #use viridis package 
  viridis::scale_fill_viridis(discrete = TRUE, name = "Median\nIncome") +
  #size the map
  coord_sf() +
  #set the theme
  theme_void()
```


<br>

##### Plot 2

*Hints:*

-   `statebins::geom_statebins()`
-   `viridis` package (`discrete = TRUE` in `scale_*` function)
-   Statebins theme

```{r}
library(statebins)


ggplot(US_income, aes(state = name, fill = income_bins)) +
  # use geom_statebins  
  geom_statebins() +
  #use viridis color scheme 
  viridis::scale_fill_viridis(discrete = TRUE, name = "Median\nIncome") +
  theme_statebins()
```



<br>

### Exercise 3

Pick any city or foreign country to build a map for. You can dress it up or make it as basic as you want. Also welcome to try building a graphic like that depicted at the end of section [6.5](https://ggplot2-book.org/maps.html#rastermaps) --- use a different region though.

```{r}
library(ggplot2)
library(dplyr)
require(maps)
require(viridis)

world_map <- map_data("world")


# select a foreign country
Germany_map <- c("Germany")

# Retrieve data about
Germany_map <- map_data("world", region = Germany_map) 


#plot state, divide by counties
ggplot(Germany_map, aes(x = long, y = lat)) +
  #title
  ggtitle("Germany") +
  #geompoly method
  geom_polygon(aes(group = group), fill = "pink" , color = "grey35") +
  #size
  coord_quickmap() +
  #theme
  theme_void()








```



<br>

## Challenge(s)

Using the `tidycensus` package and few others, try to create a map like below using [these directions](https://walkerke.github.io/tidycensus/articles/spatial-data.html). Try using a different geographical area and a different variable from the ACS.

<br>